Fully Generalized Compilation

### 2017-03-11 papers/paper.tex
Minute changes to paper
WIP on paper rough draft

% ==================================================
\section{Abstract}
% ==================================================

Assuming that two languages are \textit{isomorphic}, one language can be converted to another if given an adequate description of each language's grammar.
At an abstract level, this conversion is entirely practical. However, in practice,  
Progtran is an investigation into the feasibility of a \textbf{Fully Generalized Compiler}, a multi-language source-to-source compiler. 
Given source code in one language, a description of its grammar \textbf{A} , and a description of an output language's grammar \textbf{B}, can \textbf{A} be converted to \textbf{B} such that the output in \textbf{B} is indistinguishable from a program originally written in \textbf{B}?

@@

A \textbf{Full Generalized Compiler} is a tool for programming language researchers, providing a \textit{dynamic framework} for easily modifying syntax of input and output languages. 
A new programming language can even be fully implemented without recompiling the software. 
Progtran also uses this concept (user defined dynamic grammars) to re-implement existing source-to-source compilation tools, such as \textbf{Cython} or \textbf{2to3}.

@@

\subsection{Traditional Compilation}
Brief description of how compilers work

\subsection{Abstracting Language Definitions}
Description of grammar, transformation, and constructor files

@@

\subsection{Compilation in Glossa}
Given all of the above details, how does compilation in Glossa compare to the traditional compilation process?

(Grammar/Transform/Construction rules are read at run-time)

@@

This is not meant to be a robust benchmark. Instead, it aims to show that Cython and Transpilation yield similar results.

Output for 1000 sorting iterations:
\begin{center}
    \begin{verbatim}
        Transpile speedup:
        13.335545461765733
        Cython speedup:
        13.9866371311344
        Transpile : Cython comparison 
        0.9534
    \end{verbatim}
\end{center}

>>>>

% ==================================================
\section{Abstract}
% ==================================================
Glossa is an investigation into the feasibility of a multi-language source-to-source compiler with dynamic grammars. 
Assuming that two languages are \textit{isomorphic}, one language can be converted to another if given an adequate description of each language's grammar.
At an abstract level, this conversion is entirely practical. Given source code in one language, a description of its grammar \textbf{A} , and a description of an output language's grammar \textbf{B}, can \textbf{A} be converted to \textbf{B} such that the output in \textbf{B} is indistinguishable from a program originally written in \textbf{B}?
Once this is achieved, the grammar description for \textbf{A} can be modified, and the program will still output code in the \textbf{B} language that satisfies the above constraint.

@@

A multi-language source to source compiler is potentially a tool for programming language researchers, providing a \textit{dynamic framework} for easily modifying syntax of input and output languages. 
Glossa also uses this concept to re-implement existing source-to-source compilation tools, such as \textbf{Cython} or \textbf{2to3}.

@@

\subsection{Compilation}
Traditionally, a compiler takes a single high-level programming language, and converts it into machine instructions.
\begin{enumerate}
\item Code is lexed (seperated into tokens and tagged as keywords, punctuators, literals, etc...)
\item Tokens are parsed to build an abstract syntax tree (symbolic code representation)
\item The AST is optimized (i.e. unreachable code is removed, some loops are unrolled)
\item Machine instructions are generated.
\end{enumerate}
While some compilers and interpreters have dynamic ways to modify the syntax of a language, this functionality seems to be unneeded for most compilers.

Glossa differs only slightly from the steps above. Lexing and parsing work as above, but the rules for lexing and parsing respectively are read in at run-time. 
In the current state of the software, optimizations don't occur, but could easily be modeled as transformations to the abstract syntax tree.
Instead of generating machine instructions, Glossa generates code in another programming language. 
Instructions for doing this are also read in at run-time.
Because of its dynamic nature, it is very simple to have Glossa support multiple programming languages both for input and output.

\subsection{Abstracting Language Definitions}

As outlined above, Glossa reads in parse/generate descriptions for programming languages at run-time.
Additionally, Glossa reads in rules for transforming the abstract syntax tree.

To parse an input language, Glossa needs a description of the languages grammar, which is provided through a \textit{grammar file}.
Optional descriptions of AST transformations are provided through \textit{transformer files}.

To generate an output language from an AST, Glossa needs a set of \textit{constructor files}.

@@

@@
